<!-- README.md is generated from README.Rmd. Please edit that file -->

# `lingglosses`: Linguistic glosses and semi-automatic list of glosses creation

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(lingglosses)
```

[![CRAN version](https://www.r-pkg.org/badges/version/lingglosses)](https://cran.r-project.org/package=lingglosses)
[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](https://www.repostatus.org/badges/latest/active.svg)](https://www.repostatus.org/#active)
[![DOI](https://zenodo.org/badge/440443756.svg)](https://zenodo.org/badge/latestdoi/440443756)
[![R build status](https://github.com/agricolamz/lingglosses/workflows/R-CMD-check/badge.svg)](https://github.com/agricolamz/lingglosses/actions)

The main goal of the `lingglosses` R package is to create:

* linguistic glosses for `.html` output of `rmarkdown`;
```{r, eval = FALSE}
gloss_example(transliteration = "bur-e-ri c'in-ne-sːu",
              glosses = "fly-NPST-INF know-HAB-NEG", 
              free_translation = "I cannot fly.", 
              comment = "(lit. do not know how to)")
```

```{r, include=FALSE}
write.table(x = c("NPST", "INF", "HAB", "NEG"), 
            file = getOption("lingglosses.glosses_list"),
            row.names = FALSE, col.names = FALSE, append = TRUE)
```

* semi-automatic compiled abbreviation list.
```{r}
make_gloss_list()
```

For more details see the [html-version of the tutorial](https://agricolamz.github.io/lingglosses/).

You can also be interested in 

* [ODIN project](https://odin.linguistlist.org/);
* a Python library [`Xigt`](https://github.com/xigt/xigt);
* [scription format](https://github.com/digitallinguistics/scription) and [scription2dlx Java-script library](https://github.com/digitallinguistics/scription2dlx);
* a Python library [`pyigt`](https://github.com/cldf/pyigt).

## Installation

You can install the stable version from CRAN:

```{r, eval = FALSE}
install.packages("lingglosses")
```

You can also install the development version of `lingglosses` from [GitHub](https://github.com/agricolamz/lingglosses) with:

```{r, eval=FALSE}
# install.packages("remotes")
remotes::install_github("agricolamz/lingglosses")
```

## How to cite this package

```{r}
citation("lingglosses")
```

